---
title: "Storm Data Analysis"
author: "Kazuyuki Ishihara"
date: "2022-09-05"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

This document is a result of the analysis of "Storm Data" that is published from the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. The results suggested that "TORNADO" is the most harmful event both on population health and economic damages.

## Processing Data
### Qustion 1: Across the United States, which types of events (as indicated in the \color{red}{\verb|EVTYPE|}EVTYPE variable) are most harmful with respect to population health?  
I describe the steps of processing data as followings.  
1. read the file.  
2. extract necessary columns for subsequent data analysis.  
3. calculate the number of total health influences(variables "FATALITIES"+"INJURIES") by each event (EVTYPE).  
```{r, cache=TRUE}
Storm <- read.csv("repdata_data_StormData.csv")
library(dplyr)
StormEv <- select(Storm, EVTYPE, FATALITIES, INJURIES, PROPDMG)
EVHealth <- aggregate(FATALITIES+INJURIES~EVTYPE, data = StormEv, sum)
```
### Qustion 2: Across the United States, which types of events have the greatest economic consequences?  
4. calculate the number of total economic damages(variable "PROPDMG") by each event (EVTYPE).
```{r}
EVEcon <- aggregate(PROPDMG~EVTYPE, data = StormEv, sum)
```

## Results
### Qustion 1: Across the United States, which types of events (as indicated in the \color{red}{\verb|EVTYPE|}EVTYPE variable) are most harmful with respect to population health?  
```{r}
head(EVHealth[order(EVHealth$`FATALITIES + INJURIES`, decreasing = T),])
```
The table shows that "TORNADO" induced the maximum number of population health damages.  

### Qustion 2: Across the United States, which types of events have the greatest economic consequences?  
```{r}
head(EVEcon[order(EVEcon$PROPDMG, decreasing = T),])
```
The table shows that "TORNADO" also induced the maximum number of property damages.  

Data plot of top 6 events of each question.
```{r}
Top6Event <- data.frame("Health" = head(EVHealth[order(EVHealth$`FATALITIES + INJURIES`, decreasing = T),]), "Economy" = head(EVEcon[order(EVEcon$PROPDMG, decreasing = T),]))
library(ggplot2)
g1 <- ggplot(Top6Event, aes(Health.EVTYPE, Health.FATALITIES...INJURIES))
g1 + geom_bar(stat = "identity") + labs(x = NULL, y = "Total", title = "Damage on Population Health")
g2 <- ggplot(Top6Event, aes(Economy.EVTYPE, Economy.PROPDMG))
g2 + geom_bar(stat = "identity") + labs(x = NULL, y = "Total", title = "Damage on Property")
```